# Data Cleaning
## Import and filter data

We use data from the 2012-2022 Census of Agriculture in California. Our final sample of farms is fruit and vegetable producers (defined by primary commodity) with gross cash farm income less than \$350,000.

```{r}

library(tidyverse)
library(srvyr)

# Import 2022 data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_22.RData")

# Keep CA only
coa_22 <- coa_22$variables %>% 
  filter(state_abbr == "CA")

# Import 2007-2017 data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_071217.RData")

# Drop 2007 data
coa <- coa$variables %>% 
  filter(state_abbr == "CA" & 
           year!=2007)

# Bind rows
df <- bind_rows(coa, coa_22)

rm(coa, coa_22)

# Keep farms with less than $500,000 in sales
df <- df %>% 
  filter(tvp<500000)

# Define farms with primary commodity fruit/veg
df <- df %>% 
  filter(fruitveg==1)

```

### Define variables

**Climate smart practices**

From the 2023-24 RFA for the project, climate smart practices include (but are not limited to)

- cover cropping
- no or reduced till
- hedgerow plantings
- compost application
- prescribed grazing
- certified organic or transitioning, including organic production not certified or exempt

```{r}

# Define number of practices that is our cutoff for being defined as climate smart
cutoff <- 4

# Aggregate no and reduced till into one practice
df <- df %>% 
  mutate(
    no_or_reduced_till = case_when(
      no_till==1 | reduced_till==1 ~ 1, 
      TRUE ~ 0)
    )

# List climate smart variables
climate_smart_vars <- df %>% 
  select(
    cover_crop, no_or_reduced_till, 
    alley_cropping, manure,
    rotational_grazing, organic_farm
  ) %>% 
  names()

# Define the number of climate smart practices
df <- df %>% 
  mutate(
    climate_smart_number = rowSums(select(., all_of(climate_smart_vars))
  ))

# Define climate smart dummy variable
df <- df %>% 
  mutate(
    climate_smart = case_when(
      climate_smart_number >= cutoff ~ 1, 
      TRUE ~ 0),
    climate_smart_fct = case_when(
      climate_smart == 1 ~ "At least 4 climate smart practices", 
      climate_smart ==0 ~ "Less than 4 climate smart practices")
  )

```

### Save

```{r}

save(df,file =  "../data_processed/ca_fts.RData")

```

## Data exploration

Here we want to understand if the number of practices are associated with profit or expenses for our main sample of interest (2022, local food sales).

```{r}
library(broom)

# Define data of interest
df <- df %>%
  filter(year==2022 & local_farm==1)


# Regression profit
results <- lm(infi ~ climate_smart_number + sales_class, 
              weight = weight, 
              data = df)

tidy(results)

# Regression variable expenses
# Define list of variable expenses to loop through
varlist <- df %>% 
  select(totvarexp_alt,
         seedp_alt_var_pct, animalpurchase_alt_var_pct, 
         feed_alt_var_pct, fuelo_alt_var_pct, 
         util_alt_var_pct, mainr_alt_var_pct, 
         labor_alt_var_pct, cwork_alt_var_pct) %>% 
  names()


results <- lm(cwork_alt_var_pct ~ climate_smart_number + sales_class, 
              weight = weight, 
              data = df)

tidy(results)

```
